\documentclass[11pt]{article}
\usepackage{amsmath, amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{margin=1in}

\title{Parameterized Set-Valued Dynamical Systems: Architecture and Mathematics}
\author{Set-Valued Viz System Notes}
\date{\today}

\begin{document}
\maketitle

\section{Overview}
This document describes the parameterized custom-system pipeline for both discrete maps and continuous ODEs in the set-valued visualization tool. The goal is to allow an arbitrary number of named parameters while preserving the mathematical structure of the boundary map, normal transformation, manifold computation, and Ulam discretization.

\section{Parameter Model}
Let the parameter vector be
\begin{equation}
\mathbf{p} = (p_1, \dots, p_k) \in \mathbb{R}^k,
\end{equation}
with user-provided names (e.g., $a$, $\beta$, $\mu$). A parameter list is represented as name--value pairs and validated to satisfy:
\begin{itemize}
\item valid identifier syntax (letters, digits, underscore; starting with a letter or underscore),
\item no reserved conflicts with $x$, $y$, or built-in functions (\texttt{sin}, \texttt{cos}, \texttt{tan}, \texttt{abs}, \texttt{sqrt}, \texttt{exp}, \texttt{ln}),
\item uniqueness of names,
\item finite numeric values.
\end{itemize}
The parameter set is then injected into the evaluation context for every expression evaluation.

\section{Equation Parsing and Evaluation}
User-supplied equations are parsed once into expression trees. We use a safe expression engine and pre-process absolute values by converting \texttt{|expr|} to \texttt{abs(expr)} (including nested absolute values). Evaluation at a point $(x, y)$ proceeds by:
\begin{enumerate}
\item Clearing the evaluation context.
\item Setting $x$ and $y$.
\item Injecting all validated parameters $(p_i)$.
\item Evaluating $F_x$ and $F_y$ with the same context.
\end{enumerate}
This design makes parameter count arbitrary while keeping evaluation fast and deterministic.

\section{Discrete Map Model}
A discrete-time system is defined by expressions
\begin{equation}
(x_{n+1}, y_{n+1}) = F(x_n, y_n; \mathbf{p}),
\end{equation}
where $F = (F_x, F_y)$ is given by user equations. The set-valued map with bounded noise radius $\varepsilon \ge 0$ is
\begin{equation}
F_{\varepsilon}(x) = B_{\varepsilon}(F(x)),
\end{equation}
and the boundary map acts on the extended state $(x, n)$ with outward unit normal $n$:
\begin{equation}
E(x, n) = \left(F(x) + \varepsilon\, \hat{n}',\; \hat{n}'\right), \qquad
\hat{n}' = \frac{Df(x)^{-\top} n}{\|Df(x)^{-\top} n\|}.
\end{equation}
The Jacobian $Df(x)$ for user-defined maps is approximated by central finite differences.

\section{Continuous ODE Model}
For continuous-time flows, the vector field is
\begin{equation}
\dot{x} = f(x; \mathbf{p}), \qquad x=(x_1, x_2),
\end{equation}
where $f = (f_x, f_y)$ is specified by user equations. The boundary differential equation (BDE) on the outward unit normal bundle is
\begin{align}
\dot{x} &= f(x; \mathbf{p}) + \varepsilon\, n,\\
\dot{n} &= -Df(x; \mathbf{p})^{\top} n + \bigl\langle Df(x; \mathbf{p})^{\top} n, n \bigr\rangle n.
\end{align}
A fourth-order Runge--Kutta (RK4) integrator is used to advance both $x$ and $n$, with normalization of $n$ after each substep. When $\varepsilon = 0$, the system reduces to the standard (noise-free) ODE flow.

\section{Jacobian Approximation}
For user-defined systems, the Jacobian is computed numerically with a central difference scheme:
\begin{equation}
\frac{\partial f_i}{\partial x_j}(x) \approx
\frac{f_i(x + h e_j) - f_i(x - h e_j)}{2h},
\end{equation}
with a small step size $h = 10^{-5}$. This is used in both discrete Jacobians and continuous BDE updates.

\section{Normal Transformation}
For discrete maps, the normal update uses the inverse-transpose of the Jacobian:
\begin{equation}
\hat{n}' = \frac{Df(x)^{-\top} n}{\|Df(x)^{-\top} n\|}.
\end{equation}
For continuous flows, the BDE provides the corresponding continuous-time normal evolution. Both mechanisms preserve the unit normal constraint by explicit normalization.

\section{Ulam Discretization}
The Ulam method builds a Markov matrix by sampling a grid of boxes and propagating samples under the map (discrete) or under the time-$T$ flow (continuous). For continuous systems, $T$ is discretized into substeps $h$ (RK4), and the resulting image points are inflated by $\varepsilon$ using the same set-valued logic.

\section{Implementation Architecture}
The backend is separated into layers:
\begin{itemize}
\item \textbf{Parameter parsing/validation}: centralized in a parameter module, returning a validated parameter set.
\item \textbf{Equation parsing and evaluation}: shared user-defined equation module that parses expressions and evaluates with the parameter set.
\item \textbf{Simulation core}: discrete and continuous system implementations (map, Jacobian, BDE, RK4).
\item \textbf{WASM bindings}: thin wrappers that parse JS inputs and call core simulation logic.
\end{itemize}
This separation makes it straightforward to extend parameter sets, add new systems, and keep numerical methods stable.

\section{WASM Interface and UI Flow}
\begin{itemize}
\item The UI collects parameters as an ordered list of \texttt{\{name, value\}} rows.
\item The list is validated in the frontend (identifier syntax, duplicates, finite values).
\item The same list is sent to WASM as a JSON array, and parsed into a parameter set on the Rust side.
\item Custom map evaluation, boundary map, manifold computation, and Ulam all share the same parameter parsing pathway.
\end{itemize}
This keeps discrete and continuous custom systems symmetric and extensible.

\section{Test Coverage}
The system is validated with unit tests that cover:
\begin{itemize}
\item parameter validation (syntax, uniqueness, reserved names, finite values),
\item user-defined map evaluation and Jacobians,
\item user-defined ODE vector fields and Jacobians,
\item zero-noise limit ($\varepsilon = 0$) for BDE vs. ODE flow,
\item unit-length normal preservation in BDE integration.
\end{itemize}
These tests ensure correctness across both discrete and continuous pipelines.

\end{document}
